Theory about Normal Basis on $F(2^k)$

### 2017-03-16 dissertation/submission/intro.tex
dissertation: after submission_final revision

\section{Abstractions in Sequential Design Verification}
With the increasing size of integrated
circuits, sequential circuit designers face complicated problems of
design errors in specification models and implementations. These
errors are usually modeled as ``bad" states, and the
circuits/functional components are modeled as finite-state machines
(FSMs). Once state reachability is analyzed, the existence of errors
can be identified by checking whether ``bad" states are {\it
  reachable} from certain initial states. Temporal logic model
  checking formulations and solvers are often used for this
purpose. Once the designs and specification models are validated using
model checking, optimized implementations of sequential circuits are
synthesized. Then there is a subsequent problem that needs to be solved: how to prove
that the sequential circuit implementations are equivalent to the
original specification models? This problem is called {\it Sequential Equivalence
  Checking }(SEC) problem. When the specification is given as an arithmetic function
which canonically represents the circuit, then the problem 
becomes {\it functional verification} of sequential arithmetic circuits.

>>>>

\section{Abstractions in Sequential Design Verification}
With the increasing size of integrated
circuits, sequential circuit designers face complicated problems of
design errors in specification models and implementations. These
errors are usually modeled as ``bad" states, and the
circuits/functional components are modeled as finite-state machines
(FSMs). Once state reachability is analyzed, the existence of errors
can be identified by checking whether ``bad" states are {\it
  reachable} from certain initial states. Temporal logic model
  checking formulations and solvers are often used for this
purpose. Once the designs and specification models are validated using
model checking, optimized implementations of sequential circuits are
synthesized. Then there is a subsequent problem that needs to be solved: how to prove
that the sequential circuit implementations are equivalent to the
original specification models. This problem is called {\it Sequential Equivalence
  Checking }(SEC) problem. When the specification is given as an arithmetic function
which canonically represents the circuit, then the problem 
becomes {\it functional verification} of sequential arithmetic circuits.

### 2017-03-16 dissertation/submission/normal.tex
dissertation: after submission_final revision

%\vspace{-0.1in}
% \IncMargin{1em}
\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{Circuit polynomial ideal $J_{ckt}$, vanishing ideal $J_0$, initial
   state ideal for output $R (=0)$, inputs $\mathcal{G}(A_{init}), \mathcal{H}(B_{init})$} 
\KwOut{Circuit function $R_{final}$ after $k$ cycles of $A_{init}, B_{init}$}
  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  \tcc{$A, B$ are word-level variables, solutions to polynomial equations 
  $\mathcal{G},\mathcal{H}$ denote the initial values of preloaded operands $A_{init},B_{init}$}
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	$G \gets$GB$( \langle J_{ckt} + J_0+ from_{i-1}(R,A,B) \rangle
    )$ with ATO\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	\tcc{Using projections of varieties from abstraction theory, we get NS variables $R',A',B'$
	in terms of PS $A,B$}
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
	\tcc{By replacing NS variables with PS variables we push it to next time-frame}
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit
  verification}
\label{alg:modified}
\end{algorithm}
% \DecMargin{1em}

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=4.5in]{newfig/gf_seq_model.eps}
\caption{A typical normal basis GF sequential circuit model. $A =
  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
  $A', B', R'$ denote next-state inputs.}
\label{fig:sequential}}
\end{figure}

@@

Since the remainder from \emph{Spoly} reduction contains some bit-level variables, our objective is to compute
a polynomial that contains only word-level variables (such as $R'+\mathcal{F}(A,B)$). One possible solution to
this problem is to replace the bit-level variables by equivalent polynomials that only contain word-level
variables, e.g. $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
approach is introduced to compute corresponding $\mathcal{G}(A),\mathcal{H}(R)$ efficiently.

>>>>

@@

\begin{figure}[tbp]
\centering{
\includegraphics[width=4.5in]{newfig/gf_seq_model.eps}
\caption{A typical normal basis GF sequential circuit model. $A =
  (a_0,\dots,a_{k-1})$ and similarly $B, R$ are $k$-bit registers;
  $A', B', R'$ denote next-state inputs.}
\label{fig:sequential}}
\end{figure}

\begin{algorithm}[hbt]
\SetAlgoNoLine
\LinesNumbered
 \KwIn{Circuit polynomial ideal $J_{ckt}$, vanishing ideal $J_0$, initial
   state ideal for output $R (=0)$, inputs $\mathcal{G}(A_{init}), \mathcal{H}(B_{init})$} 
\KwOut{Circuit function $R_{final}$ after $k$ cycles of $A_{init}, B_{init}$}
  $from_0(R,A,B) = \langle R, \mathcal{G}(A_{init}), \mathcal{H}(B_{init})\rangle$\;
  \tcc{$A, B$ are word-level variables, solutions to polynomial equations 
  $\mathcal{G},\mathcal{H}$ denote the initial values of preloaded operands $A_{init},B_{init}$}
  $i = 0$\;
  \Repeat{$i == k$}
  {
  	$i \gets i + 1$\;
	$G \gets$GB$( \langle J_{ckt} + J_0+ from_{i-1}(R,A,B) \rangle
    )$ with ATO\;
	$to_i(R',A',B')\gets G\cap \mathbb F_{2^k}[R',A',B',R,A,B]$\;
	\tcc{Using projections of varieties from abstraction theory, we get NS variables $R',A',B'$
	in terms of PS $A,B$}
	$from_i \gets to_i(\{R,A,B\}\setminus \{R',A',B'\})$\;
	\tcc{By replacing NS variables with PS variables we push it to next time-frame}
  }
\Return{$from_k(R_{final})$}
\caption {Abstraction via implicit unrolling for Sequential GF circuit
  verification}
\label{alg:modified}
\end{algorithm}

@@

Since the remainder from \emph{Spoly} reduction contains some bit-level variables, our objective is to compute
a polynomial that contains only word-level variables (such as $R'+\mathcal{F}(A,B)$). One possible solution to
this problem is to replace the bit-level variables by equivalent polynomials that only contain word-level
variables, e.g., $a_i = \mathcal{G}(A), r_j = \mathcal{H}(R)$. In this section a Gaussian-elimination
approach is introduced to compute corresponding $\mathcal{G}(A),\mathcal{H}(R)$ efficiently.

### 2017-03-16 dissertation/submission/prev.tex
dissertation: after submission_final revision

The use of DDs in traditional formal verification has a lot of advantages. 
For example, DD-based model checking is very efficient as long as the DDs of sequential 
circuit can be setup. The existence of violating states in constructed DDs 
immediately deduces the violation of property. However, when the design gets
larger and larger, the time and space cost of building and storing the diagram 
increases rapidly. In our experiment of verifying a $k$-bit arithmetic circuit 
using ZDDs, when $k$ is larger than 100, the construction of ZDDs occupies 
over $99\%$ runtime of the whole procedure.
\subsection{Combinational Equivalence Checking Techniques}
The CEC problem can be solved using various methods.
Besides using canonical DDs (BDDs
\cite{BRYA86} and their word-level variants \cite{WLS}),
noncanonical representations such as And-Invert-Graph-based (AIG-based) reductions 
\cite{AIG:2002,alanmi:cec:iccad2006} are also very effective. 
Solvers for satisfiability problems (SAT) are good candidates to solve CEC problems,
as long as the miter of two circuits can be described using conjunctive normal form (CNF)
formulas. Applications of SAT on CEC include circuit-SAT solvers \cite{csat}, etc.
If the circuits being compared are structurally highly similar, AIG and circuit-SAT-based
approaches are known to be efficient.
However, when the circuits are functionally equivalent but structurally very dissimilar, none of the 
  contemporary techniques including quantifier-free bit-vector 
  (QF-BV) theory-based SMT-solvers \cite{Cryptol:fmcad09},
  offer a practical solution.

>>>>

The use of DDs in traditional formal verification has a lot of advantages. 
For example, DD-based model checking is very efficient as long as the DDs of sequential 
circuit can be setup. The existence of violating states in constructed DDs 
immediately deduces the violation of property. However, when the design gets
larger and larger, the time and space cost of building and storing the diagram 
increases rapidly. In our experiment of verifying a $k$-bit arithmetic circuit 
using ZDDs, when $k$ is larger than 100, the construction of ZDDs occupies 
over $99\%$ runtime of the whole procedure.
\subsection{Combinational Equivalence Checking Techniques}
The CEC problem can be solved using various methods.
Besides using canonical DDs (BDDs
\cite{BRYA86} and their word-level variants \cite{WLS}),
noncanonical representations such as And-Invert-Graph-based (AIG-based) reductions 
\cite{AIG:2002,alanmi:cec:iccad2006} are also very effective. 
Solvers for satisfiability problems (SAT) are good candidates to solve CEC problems,
as long as the miter of two circuits can be described using conjunctive normal form (CNF)
formulas. Applications of SAT on CEC include circuit-SAT solvers \cite{csat}, etc.
If the circuits being compared are structurally highly similar, AIG and circuit-SAT-based
approaches are known to be efficient.
However, when the circuits are functionally equivalent but structurally very dissimilar, none of the 
  contemporary techniques, including quantifier-free bit-vector 
  (QF-BV) theory-based SMT-solvers \cite{Cryptol:fmcad09},
  offer a practical solution.

### 2017-03-16 dissertation/submission/unsat.tex
dissertation: after submission_final revision

The second ``if-else" branch means: if we model check the abstracted machine $M'$ and find no
error trace, we can assert that property $p$ also holds on the original machine $M$. The reason to this assertion is that
the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an {\bf overapproximation}
of $M$, such that 
$$(M'\implies p) \implies (M\implies p)$$

@@

A naive way (and inefficient way) to identify {\it a minimal core}
using the GB computation is as follows:
Select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
({\it i.e.} if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
from the core; otherwise retain $f_i$ in $F_c$. Select a different
$f_i$ and continue until all polynomials $f_i$ are visited for
inclusion in $F_c$. This approach will produce a minimal core, as we
would have tested each polynomial $f_i$ for inclusion in the
core. This requires $O(|F|)$ calls to the GB engine, which is really
impractical.

@@

\begin{figure}[tbp]
\centering{
\includegraphics[width=4.5in]{newfig/refutation_tree.eps}}
\caption{Generating refutation trees to record UNSAT cores.}
\label{fig:refute}
\end{figure}

>>>>

The second ``if-else" branch means: if we model check the abstracted machine $M'$ and find no
error trace, we can assert that property $p$ also holds on the original machine $M$. The reason for this assertion is that
the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an {\bf overapproximation}
of $M$, such that 
$$(M'\implies p) \implies (M\implies p)$$

@@

A naive way (and inefficient way) to identify {\it a minimal core}
using the GB computation is as follows:
Select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
(i.e., if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
from the core; otherwise retain $f_i$ in $F_c$. Select a different
$f_i$ and continue until all polynomials $f_i$ are visited for
inclusion in $F_c$. This approach will produce a minimal core, as we
would have tested each polynomial $f_i$ for inclusion in the
core. This requires $O(|F|)$ calls to the GB engine, which is really
impractical.

@@

\begin{figure}[bp]
\centering{
\includegraphics[width=4.5in]{newfig/refutation_tree.eps}}
\caption{Generating refutation trees to record UNSAT cores.}
\label{fig:refute}
\end{figure}
